---
title: "simpleRegress"
author: "Nicole Orizino"
date: "Spring 2021"
output: 
        html_document:
           toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bears Regression model

### make a simple regression model - read in the data

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

This shows that the residual standard error is 5.38 on 52 degrees of freedom and the Adjusted R- squared is 0.748. 

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
newData <- data.frame(NECK=c(17,22,26))
predict(simp,newData)
predict(simp,newData,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).




```{r}
headwith <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwith)
summary.lm(headwith)
```

This shows that the residual standard error is 7.32 on 52 degrees of freedom and the Adjusted R- squared is .0532. This is less than the neck one.


```{r}
headwith <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwith)
summary.lm(headwith)
```


```{r}
headwith <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwith)
summary.lm(headwith)
```


```{r}
headwith <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwith)
summary.lm(headwith)
```