---
title: "simpleRegress"
author: "Nicole Orizino"
date: "Spring 2021"
output: 
        html_document:
           toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up Regression models

### make a simple regression model - read in the data

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```
# NECK
## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

This shows that the residual standard error is 5.38 on 52 degrees of freedom and the Adjusted R- squared is 0.748. 

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
newData <- data.frame(NECK=c(17,22,26))
predict(simp,newData)
predict(simp,newData,se.fit=TRUE)
```

This is the prediction of NECK showing it predicts 17 to be 52.76592, 22 will be 60.99353 and 26 is 67.57562.



## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

# HEADWTH
## Model LENGTH~HEADWTH

```{r}
headwith <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwith)
summary.lm(headwith)
```

This shows that the residual standard error is 7.32 on 52 degrees of freedom and the Adjusted R- squared is .0532. This is less than the neck one.


# HEADLEN
## Model LENGTH~HEADLEN

```{r}
headlenght <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenght)
summary.lm(headlenght)
```

This shows that the residual standard error is 4.24 on 52 degrees of freedom and the Adjusted R- squared is 0.8463. This shows the standard error is lower but the Adjusted R square is hihgher than the first two, meaning its better.


# WEIGHT
## Model LEGTH~WEIGHT

```{r}
weight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weight)
summary.lm(weight)
```

This shows that the residual standard error is 5.432 on 52 degrees of freedom and the Adjusted R- squared is 0.7423. The R square is less this time around as well.


# CHEST
## Model LENGTH~CHEST

```{r}
chest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chest)
summary.lm(chest)
```

This shows that the residual standard error is 4.939 on 52 degrees of freedom and the Adjusted R- squared is 0.7869. The R square is a little bigger than the last but not the biggest of the tests.




# Predictions


## New Prediction of Neck

```{r}
newData <- data.frame(NECK=c(22,17.5,17))
predict(simp,newData,se.fit=TRUE)
```

This is the new prediction of NECK showing it predicts to be  
60.99353 53.58868 52.76592.  


## Prediction of HEADLEN

```{r}
newData <- data.frame(HEADLEN=c(15.5,11.5,13))
predict(headlenght,newData,se.fit=TRUE)
```

This is the prediction of HEADLEN showing it predicts to be  
70.30793 51.94202 58.82924.  


## Prediction of HEADWTH
```{r}
newData <- data.frame(HEADWTH=c(5.5,9,8))
predict(headwith,newData,se.fit=TRUE)
```

This is the prediction of HEADWTH showing it predicts to be
55.00274 73.21694 68.01288.  


## Prediction of WEIHGHT
```{r}
newData <- data.frame(WEIGHT=c(166,105,90 ))
predict(weight,newData,se.fit=TRUE)
```

This is the prediction of WEIGHT showing it predicts as  
57.33408 52.70159 51.56245.  



## Prediction of CHEST
```{r}
newData <- data.frame(CHEST=c(35,29,28))
predict(chest,newData,se.fit=TRUE)
```

This is the prediction of CHEST showing it predicts to be  
57.94200 51.83603 50.81837.  

# Conclution

The best model based on HEADLEN as seen by adjusted R-squared, and the prediction of the length for bears 5,15,23 is   
70.31, 51.94, 58.83.  

The actual lengths of the bears was  
`r bears$LENGTH [c(5,15,23)]`.

The error in prediction for bear number 5 was about 8 inches, for bear number 15 the prediction was right on, and bear number 23 the error in prediction was just under 6 inches.



