---
title: "simpleRegress"
author: "Nicole Orizino"
date: "Spring 2021"
output: 
        html_document:
           toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bears Regression model

### make a simple regression model - read in the data

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```
# NECK
## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

This shows that the residual standard error is 5.38 on 52 degrees of freedom and the Adjusted R- squared is 0.748. 

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
newData <- data.frame(NECK=c(17,22,26))
predict(simp,newData)
predict(simp,newData,se.fit=TRUE)
```

This is the prediction of NECK showing it predicts 17 to be 52.76592, 22 will be 60.99353 and 26 is 67.57562.



## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

# HEADWTH


```{r}
headwith <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwith)
summary.lm(headwith)
```

This shows that the residual standard error is 7.32 on 52 degrees of freedom and the Adjusted R- squared is .0532. This is less than the neck one.


# HEADLEN

```{r}
headlenght <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenght)
summary.lm(headlenght)
```

This shows that the residual standard error is 4.24 on 52 degrees of freedom and the Adjusted R- squared is 0.8463. This shows the standard error is lower but the Adjusted R square is hihgher than the first two, meaning its better.


# WEIGHT

```{r}
weight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weight)
summary.lm(weight)
```

This shows that the residual standard error is 5.432 on 52 degrees of freedom and the Adjusted R- squared is 0.7423. The R square is less this time around as well.


# CHEST

```{r}
chest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chest)
summary.lm(chest)
```

This shows that the residual standard error is 4.939 on 52 degrees of freedom and the Adjusted R- squared is 0.7869. The R square is a little bigger than the last but not the biggest of the tests.


# SEX

```{r}
sex <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
abline(sex)
summary.lm(sex)
```

This shows that the residual standard error is 10.75 on 52 degrees of freedom and the Adjusted R- squared is 0.008373. The R square is extremely small.


## Predictions of HEADLEN

```{r}
newData <- data.frame(HEADLEN=c(17,13,9))
predict(headlenght,newData,se.fit=TRUE)
```

This is the prediction of NECK showing it predicts 17 to be 77.19515, 13 as 58.82924 and 9 is 40.46332 .
